Suplementary Code 4: Draw Interbirth Interval Predictons on the Last Birth Omited Data

### 2024-01-04 S5_response_to_reviewers.Rmd

---
title: 'Suplementary Code 5: Code to Respond to Reviewers'
author: "Richard Meitern"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  word_document: default
---

@@

## Fitting models

### Comment 1: Logistic regression

We can test the hypothesis that mothers are statistically more likely to stop having children after a twin birth by fitting a logistic regression model with the response variable being the binary variable *last* and the predictor variable being the binary variable *twin*.

```{r}
#fit logistic regression Estonia
logit_fit_EE <- glm(twin ~  last,
                 data = data_births_monthly_EE,
                 family = binomial(link = "logit"))
summary_logit_fit_EE <-  summary(logit_fit_EE)
summary_logit_fit_EE
```

```{r}
#fit logistic regression Other European
logit_fit <- glm(twin ~  last,
                 data = data_births_monthly,
                 family = binomial(link = "logit"))
summary_logit_fit_W <-  summary(logit_fit)
summary_logit_fit_W
```

@@

```

```{r}
#compare the precentage of twin births between last births and not last births
cat("Western dataset:\n")
tblw <- table(data_births_monthly[c("twin", "last")])
tblw
cat("\n")
#now calcualte the % by hand
cat("Last twins: ", round(100 * tblw[2,2] / sum(tblw[,2]), 2), "%\n")
cat("Prior twins: ",round(100 * tblw[2,1] / sum(tblw[,1]), 2), "%\n")

```

#### Test the Interaction

```{r}
data_births_monthly$monthly <- NULL
#data_births_monthly$pop <- "Western"
dbmFull <- rbind(data_births_monthly_EE[colnames(data_births_monthly)],
                 data_births_monthly)
dbmFull$popEE <- dbmFull$pop == "Estonia"
```

```{r}
#fit logistic regression
logit_fit_full <- glm(twin ~  last + popEE + popEE:last,
                 data = dbmFull,
                 family = binomial(link = "logit"))
summary_logit_fit_full <-  summary(logit_fit_full)
summary_logit_fit_full

```{r}
#' Get Coefficient Table
#'
#' This function returns a table with odds ratios and their confidence intervals.
#'
#' @param summary A summary object from a glm model.
#' @param datasetname A character string specifying the name of the dataset.
#'
#' @return A data frame with odds ratios and their confidence intervals.
get_coef_tbl <- function(summary, datasetname){
  tbl <- as.data.frame(summary$coefficients)
  tbl$dataset <- datasetname
  tbl$variable <- rownames(tbl)
  tbl$`Odds Ratio` <- exp(tbl$Estimate)
  tbl$`CI low` <- exp(tbl$Estimate - 1.96 * tbl$`Std. Error`)
  tbl$`CI high` <- exp(tbl$Estimate + 1.96 * tbl$`Std. Error`)
  rownames(tbl) <- NULL
  tbl <- tbl[,c(5,6,7:9)]
  return(tbl)
}

#print all coeficient tables in one table
#add dataset name to each table
coefEEtbl <- get_coef_tbl(summary_logit_fit_EE, "Estonia")
ceofWtbl <- get_coef_tbl(summary_logit_fit_W, "Western")
ceofFulltbl <- get_coef_tbl(summary_logit_fit_full, "Combined")

tbl <- rbind(coefEEtbl, ceofWtbl, ceofFulltbl)
knitr::kable(tbl, digits = 3, caption = "Odds ratios and confidence intervals for the logistic regression model with the response variable being the binary variable *twin* and the predictor variable being the binary variable *last*.")
```
```{r}
#print the frequency table
frqTbl <- aggregate(maternal_id ~ twin + last + popEE,
                    data = dbmFull, FUN = length)

# Rename the variables inside the columns
frqTbl$twin <- ifelse(frqTbl$twin == TRUE, "twins", "singleton")
frqTbl$last <- ifelse(frqTbl$last == TRUE, "Yes", "No")
frqTbl$popEE <- ifelse(frqTbl$popEE == TRUE, "Estonia", "Western")

# Rename the column names
colnames(frqTbl) <- c("Birth", "Last Breeding", "Population", "Count")
knitr::kable(frqTbl, digits = 0, caption = "Frequency table of the data used for the logistic regression model with the response variable being the binary variable *twin* and the predictor variable being the binary variable *last*.")

### 2024-01-04 S4_draw_IBI_predictions_EE_not_last.Rmd

---
title: 'Suplementary Material 4: Draw Interbirth Interval Predictons on the Last Birth Omited Data'
author: "Richard Meitern"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  word_document: default
---

>>>>

---
title: 'Suplementary Code 4: Draw Interbirth Interval Predictons on the Last Birth Omited Data'
author: "Richard Meitern"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  word_document: default
---
